---
title: "Time Series Analysis Assignment"
author: "Yashdeep"
date: "03/03/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

## Setup root directory for R Markdown

```{r setup}
knitr::opts_knit$set(root.dir = 'C:/Users/Hp/Desktop/R Programming')
```

## Environment Setup

```{r}
setwd("C:/Users/Hp/Desktop/R Programming")
getwd()
```

## Invoke Libraries

```{r}
library(dygraphs)
library(forecast)
library(TTR)
library(tseries)
library(stats)
```

## Data Import

```{r}
GasData <- forecast::gas
head(GasData)
is.ts(GasData)
ts.plot(GasData)
```

## Segmenting Time Series Data

```{r}
Actual.GasData <- window(GasData, start = c(1970,1), frequency = 12)

```

## Decomposing Time Series

```{r}
Actual.GasData.Decomp <- stl(Actual.GasData, s.window=7)
autoplot(Actual.GasData.Decomp)
```

## Train and Test Data

```{r}
Gas.Train.Data <- window(Actual.GasData, start = c(1970,1), end = c(1993,12), frequency = 12)
Gas.Test.Data <- window(Actual.GasData, start = c(1994,1), frequency = 12)
Gas.Train.Data
Gas.Test.Data
```

## Decompose Forecast

```{r}
Gas.Dec<-stl(Gas.Train.Data, s.window=7)
autoplot(Gas.Dec)
Gas.Dec.Forecast <- forecast(Gas.Dec, method="rwdrift", h=20)

Vec <- cbind(Gas.Test.Data,Gas.Dec.Forecast$mean)
ts.plot(Vec, col=c("blue", "red"), main="Actual vs Forecast")
MAPE <- mean(abs(Vec[,1]-Vec[,2])/Vec[,1])
MAPE
accuracy(Gas.Dec.Forecast$mean,Gas.Test.Data)
```

## Simple Moving Average SMA Forecast

```{r}
SMA.Gas <- SMA(Actual.GasData, n=12)
autoplot(SMA.Gas)
ts.plot(SMA.Gas,Actual.GasData, gpars = list(col = c("black", "red")))
```

## Simple Exponential Smoothing SES Forecasr

```{r}
SES.Gas <- ses(Gas.Train.Data, h = 20) 
SES.Gas
autoplot(SES.Gas)
accuracy(SES.Gas,Gas.Test.Data)
```

## Holts Winter Additive Model Forecast

```{r}
HW.Gas <- ets(Gas.Train.Data, model = "AAA")
autoplot(forecast(HW.Gas))
summary(HW.Gas)

HW.Gas.Forecast <- forecast(HW.Gas, h = 20)
accuracy(HW.Gas.Forecast, Gas.Test.Data)
```

## Holts Winter Muultiplicative Model Forecast

```{r}
HW.Gas.M <- ets(Gas.Train.Data, model = "MAM")
autoplot(forecast(HW.Gas.M))
summary(HW.Gas.M)

HW.Gas.Forecast.M <- forecast(HW.Gas.M, h = 20)
accuracy(HW.Gas.Forecast.M, Gas.Test.Data)
```

## Stationarity Test
## H0 : Time series is not Stationary
## Ha : Time series is stationary
## Accoring to dickey fuller test since the p-value is highly significant we can
## reject the null hypothesis and confidently say that time series is stationary

```{r}
adf.test(Actual.GasData)
```

## Autocorrrelations and Partial Autocorrelations

```{r}
acf(Actual.GasData, lag.max = 90)
acf(Gas.Train.Data, lag.max = 90)
pacf(Gas.Train.Data, lag.max = 90)
```

## Manual ARIMA Model

```{r}
Gas.Arima <- arima(Gas.Train.Data, order =  c(3,0,2))
Gas.Arima
hist(Gas.Arima$residuals, col = c("beige"))
tsdisplay(residuals(Gas.Arima), lag.max=15, main='Model Residuals')
```

## H0: Residuals are independent
## Ha: Residuals are not independent
## Residuals are independent

```{r}
Box.test(Gas.Arima$residuals)
```

## Forecast using Manual ARIMA Model

```{r}
Arima.Gas.Forecast <- forecast(Gas.Arima,h=20)
Arima.Gas.Forecast
autoplot(Arima.Gas.Forecast)
accuracy(Arima.Gas.Forecast,Gas.Test.Data)
```

## Auto ARIMA Model

```{r}
Gas.Auto.Arima = auto.arima(Gas.Train.Data, seasonal = TRUE)
hist(Gas.Auto.Arima$residuals)
Gas.Auto.Arima
```

## Forecast Auto ARIMA Model

```{r}
Gas.Forecast.AArima <- forecast(Gas.Auto.Arima, h = 20)
autoplot(Gas.Forecast.AArima)
accuracy(Gas.Forecast.AArima,Gas.Test.Data)
```

